### 2020-04-15 Report/aws_325ghz_study.tex

\section{Data and methods}

@@

\subsection{Sensor characteristics}
%
The AWS satellite was assumed to be at an altitude of 600\,km. Data were
generated for sensor viewing angles (the angle from nadir) from 0$^\circ$ to
45$^\circ$, in steps of 5$^\circ$.

AWS channels that were included, and considered as fixed in terms of positions
and widths, are found in Table\dots [* Finish after talking to Anders at
Omnisys. * ]

[* Clarify what polarisation we assume for the different channels *]

[* Inderpreet: Add equation and text about polarisation rotation (ie equations
in the Sow) *]

[* Tbs, transmissivities and Jacobians are averaged over the channels *]

[* Describe settings for ATMS together with results, if included *)

@@

\subsection{Overview of the 183, 229 and 325\,GHz bands}
%
The simulations in this section are based on five Fascod climate scenarios
(Sec.~\ref{sec:fascod}), and performed with a fixed sensor viewing angle of
25$^\circ$. No hydrometeors are included. For simpler interpretation of the
results, the reflectivity is set to have no frequency variation, i.e.\ the same
value is applied for all bands. The reflective is varied between 0 and 0.5.

In the figures covering transmissivities and Jacobians, ICI's 325\,GHz channels
are applied (Table~?).

@@

\begin{figure}[p]
  \centering
  \includegraphics[width=0.75\textwidth]{fascod_tb_r000}
  \caption{Solid lines show brightness temperatures for the lower wing of the
    183.31\,GHz transition. Dashed lines represent double-sideband measurements
    around 325.15\,GHz, with the mean of upper and lower band shown on the low
    side. Dash-dotted lines represent double-sideband measurements around
    229.00\,GHz in the same manner, but only over the bandwidth of the
    potential channel. For all three bands, spectra for all five
    Fascod scenarios are included.}
  \label{fig:tb:r000}
\end{figure}
\begin{figure}[p]
  \centering
  \includegraphics[width=0.75\textwidth]{fascod_tb_r050}
  \caption{Same as figure above, but with a surface reflectivity of 0.5.}
  \label{fig:tb:r050}
\end{figure}
\begin{figure}[p]
  \centering
  \includegraphics[width=0.75\textwidth]{fascod_tr_mls}
  \caption{Transmissivity to space as a function of altitude, for the
    mid-latitude summer scenario. Solid lines represent 183\,GHz channels,
    dashed lines 325\,GHz channels (ICI-5/6/7) and dashed-dotted the 229\,GHz
    channel. These transmissivities are independent of surface reflectivity.}
  \label{fig:tr:mls}
\end{figure}
\begin{figure}[p]
  \centering
  \includegraphics[width=0.75\textwidth]{fascod_wf_mls_025}
  \caption{Jacobians for relative humidity. Line symbols as in Fig.~\ref{fig:tr:mls}.
    Surface reflectivity is set to 0.25.}
  \label{fig:wf:mls:025}
\end{figure}

@@

The altitude variation of transmissivity exhibits same basic shape between all
the 183 and 325\,GHz channels, but the altitude of optical thickness of 1
(transmissivity of $e^{-1}\approx0.37$) varies. The curves of the end 183\,GHz
channels brackets the ones corresponding to the ICI 325\,GHZ channels. The later
channels show a higher deviation from 1 at 150\,hPa, due to some (weak) ozone
attenuation in the stratosphere. The 229\,GHz has the highest transmissivities.

@@

The basic shape of the Jacobian is similar for all 183 and 325\,GHz channels,
but the peak altitude varies. The end 183\,GHz channels brackets the 325\,GHz
ICI ones, as for transmissivity. These Jacobians are negative, meaning that a
change towards higher humidity results in decreased brightness temperatures.
The 229\,GHz Jacobian is positive close to the surface, as this channel has
sensitivity all the way down to the surface and more water vapour just above a
reflecting surface results in higher brightness temperatures. The switch from
negative to positive Jacobian depends on atmospheric scenario and surface
reflectivity. For dry conditions also the outer 183 and 325\,GHz channels can
have a positive part, as long as the reflectivity is not close to 0.

>>>>

\section{Assumptions on AWS characteristics}
%
The assumption listed below are primarily taken for the Statement of Work
(SoW), complemented with information provided by Anders Emrich at Omnisys
(private communication, April 14 2020). The later data represent what Omnisys
considers to be technically feasible inside the budget of the project.

\subsection{Channel specifications}
%
Channels around 89, 166, 183 and 229\,GHz are assumed to be fixed, with name,
position and widths defined in Table~\ref{tab:fixed:chs}. AWS channels 31 to 36
will be implemented as a single front-end, but are here still considered as two
``bands'', denoted as 166 and 183\,GHz.

For 325\,GHz, it will be assumed that channels can be placed between 0.75 and
8.0\,GHz in terms of intermediate frequency (IF), and that up to 4 channels
inside this range can be implemented. There should be some gap between these
channels, above 100\,-\,200\,MHz, to facilitate the technical implementation.
For comparison, Table~\ref{tab:fixed:chs} includes also the position of the ICI
325\,GHz channels. As the ICI channels extend to 11\,GHz in IF, they will not
be considered for AWS but will be applied for demonstration and reference
purposes.

For simplicity, the frequency variation of all channel responses is assumed to
be rectangular. This selection shall not be interpreted as any opinion on the
optimal shape of these responses; a significant deviation from a rectangular
shape should be of no concern as long as the actual shape is well
characterised.

\begin{table}[!bt]
  \begin{minipage}[b]{0.5\linewidth}
  \centering  
  \begin{tabular}[c]{c|c|c}
    Channel & Frequency   & Bandwidth \\
    name    & [GHz] &  [MHz] \\
    \hline
    AWS-21  & \phantom{0}89.000 & 4000\\
    AWS-31  & 165.500 & 2800\\
    AWS-32  & 176.311 & 2000\\
    AWS-33  & 178.811 & 2000\\
    AWS-34  & 180.311 & 1000\\
    AWS-35  & 181.511 & 1000\\
    AWS-36  & 182.311 & \phantom{0}500\\
    AWS-41  & 229.000 & 2000\\
    \hline
  \end{tabular}
  \end{minipage}%
  \begin{minipage}[b]{0.5\linewidth}
  \centering  
  \begin{tabular}[c]{c|c|c}
    Channel & Frequency   & Bandwidth \\
    name    & [GHz] &  [MHz] \\
    \hline
    ICI-5  & 325.15$\pm$9.50 & 3000\\
    ICI-6  & 325.15$\pm$3.50 & 2400\\
    ICI-7  & 325.15$\pm$1.50 & 1600\\
    \hline
  \end{tabular}
  \end{minipage}  
  \caption{Fixed channel position and widths. The left table covers channels of
    AWS, with values taken from SoW. Channels AWS-21 to AWS-36 are at this
    moment mandatory for AWS, while AWS-41 is so far tentative. The right table
    lists the values of some channels of ICI \citep{eriksson:towar:20}.}
  \label{tab:fixed:chs}
\end{table}

\subsection{Noise}
%
The nominal reciever noise temperature, $T_r$ of each band is listed in
Table~\ref{tab:trec}. The noise of data for individual channels is calculated
as
\begin{equation}
  \sigma^i = c \frac{T^i_r+T^i_a}{\sqrt{\Delta\!f^i\Delta t}}
\end{equation}
where $\sigma^i$ is the noise standard deviation of channel $i$, $c$ is a
scaling factor (see below), $T^i_r$ is the reciever noise temperature of the
channel, $T^i_a$ is the antenna temperature of the channel, $\Delta\!f^i$ is the
bandwidth of channel and $\Delta t$ is the integration time.

The factor $c$ is set to 1.2 to incorporate the additional noise caused by the
calibration process. The bandwith applied is the one found in columns 3 of
Table~\ref{tab:fixed:chs}. The integration time is set to match an effective
15\,km resolution at 183\,GHz. This time was estimated to 3\,ms.

\begin{table}[!bt]
  \centering
  \begin{tabular}[b]{c|ccccc}
    Band [GHz]   & 89 & 166 & 183 & 229 & 325 \\
    \hline
    $T_r$ [K]    & 390 & 650 & 650 & {\it 1000}  & 1200 \\
    \hline
  \end{tabular}
  \caption{Assumed reciever noise temperatures. The value for 229\,GHz is
    selected by us, remaining values are present assumptions at Omnisys.}
  \label{tab:trec}
\end{table}

\subsection{Various}
%
The AWS satellite is assumed to be at an altitude of 600\,km. Data were
generated for sensor viewing angles (the angle from nadir) from 0$^\circ$ to
45$^\circ$, in steps of 5$^\circ$.

The polarisation response of the bands is not yet determined and for simplicity
``quasi-horisontal'' (defined in same way as for e.g.\ MWS) is throughout
assumed. Antenna temperatures are obtained as:
\begin{equation}
  \label{eq:polrot}
  T_a = T_b^V\sin^2(\theta) + T_b^H\cos^2(\theta)
\end{equation}
where $T_b^V$/$T_b^H$ is the brightness temprature at vertical/horisontal
polarisation, respectively, and $\theta$ is the viewing angle (from nadir).

\section{Input data and software}

@@

@@

\subsection{Overview of the 183, 229 and 325\,GHz bands}
%
The simulations in this section are based on five Fascod climate scenarios
(Sec.~\ref{sec:fascod}), and performed with a fixed sensor viewing angle of
25$^\circ$. No hydrometeors are included. For simpler interpretation of the
results, the reflectivity is set to have no frequency variation, i.e.\ the same
value is applied for all bands. The reflective is varied between 0 and 0.5.
In the figures covering transmissivities and Jacobians, the specifications
of ICI is applied for 325\,GHz channels (Table~\ref{tab:fixed:chs}).

@@

\begin{figure}[p]
  \centering
  \includegraphics[width=0.75\textwidth]{fascod_tb_r000}
  \caption{Solid lines show brightness temperatures for the lower wing of the
    183.31\,GHz transition. Dashed lines represent double-sideband measurements
    around 325.15\,GHz, with the mean of upper and lower band shown on the low
    side. Dash-dotted lines represent double-sideband measurements around
    229.00\,GHz in the same manner, but only over the bandwidth of the
    potential channel. For all three bands, spectra for all five
    Fascod scenarios are included.}
  \label{fig:tb:r000}
\end{figure}
\begin{figure}[p]
  \centering
  \includegraphics[width=0.75\textwidth]{fascod_tb_r050}
  \caption{Same as figure above, but with a surface reflectivity of 0.5.}
  \label{fig:tb:r050}
\end{figure}
\begin{figure}[p]
  \centering
  \includegraphics[width=0.75\textwidth]{fascod_tr_mls}
  \caption{Transmissivity to space as a function of altitude, for the
    mid-latitude summer scenario. Solid lines represent 183\,GHz channels,
    dashed lines 325\,GHz channels and dashed-dotted the 229\,GHz channel
    (Table~\ref{tab:fixed:chs}). These transmissivities are independent of
    surface reflectivity.}
  \label{fig:tr:mls}
\end{figure}
\begin{figure}[p]
  \centering
  \includegraphics[width=0.75\textwidth]{fascod_wf_mls_025}
  \caption{Jacobians for relative humidity. Line symbols as in Fig.~\ref{fig:tr:mls}.
    Surface reflectivity is set to 0.25.}
  \label{fig:wf:mls:025}
\end{figure}

@@

The altitude variation of transmissivity exhibits same basic shape between all
the 183 and 325\,GHz channels, but the altitude of optical thickness of 1
(transmissivity of $e^{-1}\approx0.37$) varies. The curves of the edge 183\,GHz
channels (AWS-32 and 36) brackets the ones corresponding to the ICI 325\,GHZ
channels. The later channels show a higher deviation from 1 at 150\,hPa, due to
some (weak) ozone attenuation in the stratosphere. The 229\,GHz has the highest
transmissivities.

@@

The basic shape of the Jacobian is similar for all 183 and 325\,GHz channels,
but the peak altitude varies. Again the edge 183\,GHz channels brackets the
325\,GHz ICI ones, as for transmissivity. These Jacobians are negative, meaning
that a change towards higher humidity results in decreased brightness
temperatures. The 229\,GHz Jacobian is positive close to the surface, as this
channel has sensitivity all the way down to the surface and more water vapour
just above a reflecting surface results in higher brightness temperatures. The
switch from negative to positive Jacobian depends on atmospheric scenario and
surface reflectivity. For dry conditions also the outer 183 and 325\,GHz
channels can have a positive part, as long as the reflectivity is not close to
0.

\section{Notes}
\begin{itemize}
\item Discuss selection of polarisation
\end{itemize}
